State and Parameter Estimation for Spatio-Temporal Neural Fields

### 2010-05-06 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
the plot of MRMSE is added

\subsection{State and Parameter Estimates} To demonstrate the performance of the parameter and state estimation, $200$ realisations of the data were generated. Each realisation consisted of 200 ms of data and the estimation was applied to the final 200 ms, allowing the model to stabilise. The initial state and parameters were unknown to the estimator. The state estimates were evaluated by comparing the field reconstruction to the true field using the mean over time of the RMSE and standard deviation of the RMSE. The distribution of the parameters and the resulting kernel are shown in figure~\ref{fig:Parameters} and figure~\ref{fig:KernelEstimates} respectively. Although a wide distribution can be seen over the parameter estimates the ratio between them remans in a narrow range as shown in figure~\ref{fig:ParametersRatio}. One snapshot for the original and the field estimate and the corresponding error are illustrated in figure~\ref{fig:FieldEstimate}.
\begin{figure}
\includegraphics[width=0.5\textwidth]{./Graph/KernelEstimate_KernelFreqResponse.pdf}
\caption{Mean kernel estimates over 272 runs ;  true and estimated kernels are shown by solid
and dashed line respectively, the red area shows 95\% confedence interval; top: spatial domain; bottom: frequency domain}
\label{fig:KernelEstimates}
\end{figure}
\begin{figure}[th]
\subfigure[][]{\includegraphics[width=0.24\textwidth]{./Graph/theta0_theta1_ratio.pdf}}
\subfigure[][]{\includegraphics[width=0.24\textwidth]{./Graph/theta0_theta2_ratio.pdf}}\\
\caption{Distribution of the parameters ratios over 272 runs.}
\label{fig:ParametersRatio}
\end{figure}
\begin{figure*}[!th] 
\subfigure[][]{\includegraphics[width=2in]{./Graph/TrueField90.pdf}}
\subfigure[][]{\includegraphics[width=2in]{./Graph/EstimatedField90.pdf}}
\subfigure[][]{\includegraphics[width=2in]{./Graph/EstimationError90.pdf}}\\
\caption{One snapshot of the spatial field; a: True field; b: Estimated field; c: The absolute error.}
\label{fig:FieldEstimate}
\end{figure*}
\section{Discussion}\label{DiscussionSection}
\begin{itemize}
	\item if we don't know the frequency analysis or the system is marginally sampled or under-sampled, the best we can do is have the basis functions the same as the sensor kernel. \wtf{why wouldn't it be better to have, say, twice as many basis functions as we have sensors, at half the width?} 
	\item alternatives to frequency analysis are model selection... \wtf{This frequency analysis \emph{is} model selection. Maybe what's meant here is that other model selection approaches could be applied? Some sort of IC optimisation?} 
	\item the width of the connectivity basis functions \wtf{do you mean width of the kernel? I'm not sure about this - as the parameters increase the kernel width increases right? So maybe you mean the basis functions, but don't we estimate these widths from the frequency analysis?} here is known exactly. In practice we don't 
	\item the number of basis functions could be set higher to get a better representation of the field, however this is limited by computation complexity. \wtf{Unless you are going to specify the computational complexity, in terms of the number of basis functions, it might be better to say something like ``there is a relationship between the number of basis functions that represent the field and the computational complexity of the estimation algorithm - the more basis functions the longer it takes to estimate''} 
	\item Discuss the novelty of using the reduced model (BrainIDE) for creating patient specific models. \omg{hell yes. This should be front and centre!} 
	\item The size of the dentritic arbour is consistent in the cortex. \omg{?} 
	\item Spatial frequency of field and the relationship to the required density of electrodes/sensors required for estimation or to capture the required information without aliasing. 
	\item The flexibility of the approach 
	\item Combining this method with other to incorporate longer range hetrogeneous connectivity. \omg{awesome} 
	\item The need for activity to perform estimation. \omg{To Learn You Must Perturb!} 
	\item Recent studies using neural field models have demonstrated the theoretical implications of specific connectivity structure of the cortical columns in neural field models, where the connectivity kernel governs the bifurcation points of the system \cite{Hutt2005} and types of oscillations that can be generated \cite{Schmidt2009}. This implies that if we could estimate the connectivity structure for an individual, then we could capture patient specific neurodynamics that lead to various oscillatory states.
	\item Estimating functional cortical connectivity via EEG promises to shed light on the hierarchy of brain regions that are involved in specific tasks, motivating the use of techniques like DCM. Other comparable methods are based on auto-regressive (AR) modelling of the EEG, MEG, or fMRI times series \todo{[refs!]}. Studies have concentrated on finding functional connectivity patterns using information contained in the AR coefficients using techniques like Granger causality \cite{Hesse2003}, the direct transfer function \cite{Kaminski1991}, and partial directed coherence \cite{Sameshima1999}. Again, like the DCM approach, these methods estimate long range connectivity patterns that does not provide a clear relationship between the continuum field models and data.
	\item Long range connectivity can be estimated using DTI \cite{Knock2009}, however this short range connectivity can not estimated using DTI
\end{itemize}

@@

	Consider the zero-mean process $e_t\left(\mathbf r\right)$ with covariance as defined by \ref{eq:FieldCovarianceFunction} then 

>>>>

\subsection{State and Parameter Estimates} To demonstrate the performance of the parameter and state estimation, $200$ realisations of the data were generated. Each realisation consisted of 200 ms of data and the estimation was applied to the final 200 ms, allowing the model to stabilise. The initial state and parameters were unknown to the estimator. The distribution of the parameters and the resulting kernel are shown in figure~\ref{fig:Parameters} and figure~\ref{fig:KernelEstimates} respectively. Although a wide distribution can be seen over the parameter estimates the ratio between them remans in a narrow range as shown in figure~\ref{fig:ParametersRatio}. The state estimates for one realization were evaluated by comparing the field reconstruction to the true field using the mean over time of the Root Mean Square Error (RMSE) and standard deviation of the RMSE giving $0.19\pm 0.01$. The mean of RMSE over time is also shown in figure~\ref{fig:MRMSE} One snapshot for the original and the field estimate and their corresponding error are illustrated in figure~\ref{fig:FieldEstimate}.
\begin{figure}
\includegraphics[width=0.5\textwidth]{./Graph/KernelEstimate_KernelFreqResponse.pdf}
\caption{Mean kernel estimates over 200 runs ;  true and estimated kernels are shown by solid
and dashed line respectively, the red area shows 95\% confedence interval; top: spatial domain; bottom: frequency domain}
\label{fig:KernelEstimates}
\end{figure}
\begin{figure}[th]
\subfigure[][]{\includegraphics[width=0.24\textwidth]{./Graph/theta0_theta1_ratio.pdf}}
\subfigure[][]{\includegraphics[width=0.24\textwidth]{./Graph/theta0_theta2_ratio.pdf}}\\
\caption{Distribution of the parameters ratios over 272 runs.}
\label{fig:ParametersRatio}
\end{figure}
  \begin{figure}\label{fig:MRMSE}
   	\begin{center}
   		\includegraphics[width=0.5\textwidth]{./Graph/MRMSE.pdf} 
   	\end{center}
   	\caption{The mean of RMSE over time.} 
   \end{figure}

\begin{figure*}[!th] 
\subfigure[][]{\includegraphics[width=2in]{./Graph/TrueField90.pdf}}
\subfigure[][]{\includegraphics[width=2in]{./Graph/EstimatedField90.pdf}}
\subfigure[][]{\includegraphics[width=2in]{./Graph/EstimationError90.pdf}}\\
\caption{One snapshot of the spatial field; a: True field; b: Estimated field; c: The absolute error.}
\label{fig:FieldEstimate}
\end{figure*}
\section{Discussion}\label{DiscussionSection}
\begin{itemize}
	\item if we don't know the frequency analysis or the system is marginally sampled or under-sampled, the best we can do is have the basis functions the same as the sensor kernel. \wtf{why wouldn't it be better to have, say, twice as many basis functions as we have sensors, at half the width?} 
	\item alternatives to frequency analysis are model selection... \wtf{This frequency analysis \emph{is} model selection. Maybe what's meant here is that other model selection approaches could be applied? Some sort of IC optimisation?} 
	\item the width of the connectivity basis functions \wtf{do you mean width of the kernel? I'm not sure about this - as the parameters increase the kernel width increases right? So maybe you mean the basis functions, but don't we estimate these widths from the frequency analysis?} here is known exactly. In practice we don't 
	\item the number of basis functions could be set higher to get a better representation of the field, however this is limited by computation complexity. \wtf{Unless you are going to specify the computational complexity, in terms of the number of basis functions, it might be better to say something like ``there is a relationship between the number of basis functions that represent the field and the computational complexity of the estimation algorithm - the more basis functions the longer it takes to estimate''} 
	\item Discuss the novelty of using the reduced model (BrainIDE) for creating patient specific models. \omg{hell yes. This should be front and centre!} 
	\item The size of the dentritic arbour is consistent in the cortex. \omg{?} 
	\item Spatial frequency of field and the relationship to the required density of electrodes/sensors required for estimation or to capture the required information without aliasing. 
	\item The flexibility of the approach 
	\item Combining this method with other to incorporate longer range hetrogeneous connectivity. \omg{awesome} 
	\item The need for activity to perform estimation. \omg{To Learn You Must Perturb!} 
	\item Recent studies using neural field models have demonstrated the theoretical implications of specific connectivity structure of the cortical columns in neural field models, where the connectivity kernel governs the bifurcation points of the system \cite{Hutt2005} and types of oscillations that can be generated \cite{Schmidt2009}. This implies that if we could estimate the connectivity structure for an individual, then we could capture patient specific neurodynamics that lead to various oscillatory states.
	\item Estimating functional cortical connectivity via EEG promises to shed light on the hierarchy of brain regions that are involved in specific tasks, motivating the use of techniques like DCM. Other comparable methods are based on auto-regressive (AR) modelling of the EEG, MEG, or fMRI times series \todo{[refs!]}. Studies have concentrated on finding functional connectivity patterns using information contained in the AR coefficients using techniques like Granger causality \cite{Hesse2003}, the direct transfer function \cite{Kaminski1991}, and partial directed coherence \cite{Sameshima1999}. Again, like the DCM approach, these methods estimate long range connectivity patterns that does not provide a clear relationship between the continuum field models and data.
	\item Long range connectivity can be estimated using DTI \cite{Knock2009}, however this short range connectivity can not estimated using DTI
\end{itemize}

@@

	Consider the i.i.d. noise term $e_t\sim\mathcal{GP}(\mathbf 0,\gamma(\mathbf{s}-\mathbf{r}))$ then 
